Improved Bounds for Online Balanced Repartitioning

### 2020-04-22 paper/disc20clustering.tex
Disc not waoa
Remove: perfect partition reachable within one swap

% for my small screen
\setlength{\textwidth}{8cm}

@@

@@

\subsection{Perfect Partition reachable within one swap}

\maciek{Best to include it in the improved LB omega(lk) to help increase the lb by k-1}

  (The version for two clusters, but extensions possible.)
  Results: LB = UB = k+1.

  In this section we show that the competitive ratio of the BRP variant, where the perfect partition is reachable within one swap is exactly k+1.

  Previously the best upper bound was $O(k^2)$ of DET \cite{repartition-disc}, and the best lower bound was $k-1$, utilizing the reduction from paging from \cite{repartition-disc}.
  \maciek{It is possible to improve the reduction, i.e., to reduce from paging with cache $k$, but it is impossible to have reduction from paging with cache $k+1$.}

  \begin{theorem}
    If the perfect partition is reachable within one swap, the cost of any deterministic algorithm is at least k+1.
  \end{theorem}

  \begin{proof}
    \maciek{pre-sketch}
    A request to any pair. The algorithm puts it on the wrong cluster, evicting some node and paying $1$.
    We continue to issue the requests to the evicted node (it changes every time), alg pays $k-2$.
    We issue the final request, and alg must move the initial pair, paying $2$.
  \end{proof}

  \begin{theorem}
    There exists a (k+1)-competitive algorithm for the BRP variant, where the perfect partition is reachable within one swap.
  \end{theorem}

  \begin{proof}
    \maciek{Maybe PPL can be used instead?}

    \maciek{early sketch}
    The algorithm maintains components of nodes, and never splits them.
    Intially all nodes are in singleton components.
    The only choice is where to place the component, and possibly which components to evict.

  The algorithm places the first component, denoted $Z$, on an arbitrary cluster, and evicts an arbitrary node (it is a singleton).
  Consider a request $(a,b)$.
  If $a \in Z$ (also symmetry $b \in Z$), then we move the entire component $Z$ to the cluster $I(b)$, evicting a singleton node that is missplaced (i.e., is not on its initial cluster), and an arbitrary node (we later show that there exists a node that is misplaced).
  Otherwise, $I(a) = I(b)$, and we place the component there, evicting an arbitrary node (non-missplaced?).

  If a component collocated with $Z$ grows to $k-1$, we evict $Z$ and terminate.
  \end{proof}

\subsection{Lower bound $2k-1$}

\maciek{Dont include in WAOA}

We show the reduction from the paging problem with the universe $U$ of size $2k$ and the cache of size $2k-1$.
Recall that the lower bound for the competitive ratio for such paging problem is $2k-1$.

Similarly to the lower bound TODO ref $\Omega(kl)$, we construct components of nodes, and issue the requests if the algorithm splits the component.
In the proof in this section, the components produced are oblivious to the configuration of the algorithm, yet the requests issued depend on if the algorithm splits the components. In $\Omega(kl)$ LB, the components were constructed upon the actions of the algorithm, too.

\maciek{TODO: figure of three states of OPT. It should fit next to the other lower bound figure?}

\begin{theorem}
  %No deterministic algorithm for BRP with two clusters can achieve a competitive ratio smaller than $2k-1$.

  Fix any $k$. If there exists a $\gamma$-competitive deterministic algorithm $B$ for Dynamic Bisection Problem with cluster size $k$, then there exists a $\gamma$-competitive deterministic algorithm $P$ for the paging problem with cache size $2k-2$ and where the number of different pages is $2k-1$. 
\end{theorem}

\begin{proof}
  The pages are denoted by $p_1,p_2,\ldots,p_{2k-1}$. Without loss of generality, we assume that the initial cache is equal to $p_1,p_2,\ldots,p_{k-2}$.
  We fix any input sequence $\sigma^P = (\sigma^P_1, \sigma^P_2, \sigma^P_3, \ldots)$ for the caching problem, where $\sigma^P_t$ denotes the $t$-th accessed page.
  We show how to construct an online algorithm $P$ for the caching problem that proceeds in the following online manner.
  The algorithm internally runs the algorithm~$B$, starting on the initial assignment of nodes to clusters that will be defined below.
  For a requested page $\sigma^P_t$, it creates a subsequence of~communication requests for the BRP problem, runs $B$ on them, and serves $\sigma^P_t$ on the basis of $B$'s responses.

  We use $2k$ nodes for the BRP problem: $2k-1$ caching nodes denoted $p_1,p_2, \ldots, p_{2k-1}$, and an auxiliary node $a$.
  We partition nodes into sets $A = \{ a, p_1 \}$, $B = \{ p_2, p_3, \ldots, p_{k+1}\}$, and $C = \{ p_{k+2}, p_{k+3} \ldots, p_{2k-1}\}$.
  Their cardinalities are $|A| = 2$ and $|B| = |C| = k-1$.

  We say that the node configuration is \emph{well-aligned} if any of the following conditions hold:
  \begin{enumerate}
    \item One cluster contains $B$ and one node from $A$ and the other cluster contains $C$ and the other node from $A$.
    \item One cluster contains $B$ and one node from $C$ and the other cluster contains $A$ and $k-2$ nodes from set $C$.
    \item (symmetric to the previous condition) One cluster contains $C$ and one node from $B$ and the other cluster contains $A$ and $k-2$ nodes from set $B$.
  \end{enumerate}
  We show a bijection between cache contents and well-aligned configurations.
  We say that a caching node $p_i \in S$, where $S \in \{ A, B, C\}$ belongs to the cache if the entire set $S$ is collocated in a cluster.
  Otherwise, if $p_i$ is located on a different cluster than the other nodes from $S$, we say that it is outside of the cache.
  (Without loss of generality, we may assume that the cache of any caching algorithm is always full, i.e., consists of $k-1$ pages.)
  If the configuration $c$ of a~BRP algorithm is well-aligned, $\textsc{cache}(c)$ denotes the corresponding cache contents.
  The initial configuration for the BRP problem is the well-aligned configuration corresponding to the initial cache (pages $p_1,p_2,\ldots,p_{2k-2}$ in the cache), with sets $A$ and $B$ non-split and the node $p_{2k-1}$ from the set $C$ split.

When a request $p$ to paging arrives, we create a subsequence $\comm(p)$ of communication requests for the BRP problem.
We identify the corresponding caching node $p$ and the set $S \in \{ A,B,C \}$ that contains it.
The construction of $\comm(p)$ depends on the actions of $B$.
To abstract from the actions of $B$ in the construction, we use the notion of the adversary components.
Forming an adversary component means that the adversary continuously issues requests to any two nodes of a component that are split.
The request sequence $\comm(p)$ consists of the requests the adversary issues for $B$ after forming an adversary component $S$.
When $B$ collocates all nodes of $S$, the adversary dismisses the adversary component $S$, i.e., no longer issues requests if $B$ splits the nodes of $S$.

Note that $B$ must eventually achieve such a node configuration: otherwise its cost would be arbitrarily large while a sequence of repeated $\comm(\sigma^P_t)$ subsequences can be served at~a~constant cost.
Therefore, the competitive ratio of~$B$ would then be unbounded.
We denote the resulting sequence of $\comm(\sigma^P_t)$ subsequences by $\comm_t(\sigma^P_t)$.

To construct the response to the caching request $\sigma^P_t$, the algorithm $P$ runs $B$ on $\comm_t(\sigma^P_t)$.
Right after processing $\comm_t(\sigma^P_t)$, the node configuration $c$ of $B$ is well-aligned and $\sigma^P_t$ is collocated with other nodes from $S_t$ where $S_t \in \{A,B, C\}$ and $\sigma^P_t \in S_t$.
Hence, $P$ may change its cache configuration to $\textsc{cache}(c)$: such a response is feasible, because since $S_t$ is collocated, it is included by $P$ in the cache.
Furthermore, we may relate the cost of $P$ to the cost of~$B$: If $P$ modifies the cache contents, the corresponding cost is $1$, as exactly one page has to
be fetched.
Such a change occurs only if $B$ changed the configuration (at a~cost of at least $2 \cdot \alpha$).
Therefore, $2\cdot \alpha \cdot P(\sigma^P_t) \leq B(\comm_t(\sigma^P_t))$, which, summed over all requests from sequence $\sigma^P$, yields $2 \cdot
\alpha \cdot P(\sigma^P) \leq B(\sigma^B)$.

Now we show that there exists an (offline) solution \OFF to $\sigma^B$, whose cost is exactly ${2 \cdot \alpha \cdot \OPT(\sigma^P)}$.
Recall that, for a caching request $\sigma^P_t$, $\sigma^B$ contains the~corresponding sequence $\comm_t(\sigma^P_t)$.
Before serving the first request of~$\comm_t(\sigma^P_t)$, \OFF changes its state to a well-aligned configuration corresponding to the cache of $\OPT$ right after serving caching request $\sigma^P_t$.
This ensures that the subsequence $\comm_t(\sigma^P_t)$ is free for \OFF.

Now, we claim that the cost of node migration of \OFF is $2 \cdot \alpha$ (two caching nodes are swapped) if $\OPT$ performs a fetch, and 0 if
$\OPT$ does not change its cache contents.
Within each type of configuration with $S \in \{ B, C\}$ split, $\OPT$ may achieve a configuration with a different node from $S$ outside of the cache by exchanging the split node with any node from $S$.
To transition between any of three configuration types ((1) with $A$ (2) with $B$ split and (3) with C split), it suffices to perform a single exchange.
Therefore, $\OFF(\comm_t(\sigma^P_t)) = 2 \cdot \alpha \cdot \OPT(\sigma^P_t)$, which summed over the entire sequence $\sigma^P$ 5yields $\OFF(\sigma^B) = 2 \cdot \alpha \cdot \OPT(\sigma^P)$.

As $B$ is $\gamma$-competitive for the BRP problem, there exists a constant $\beta$, such that for any sequence $\sigma^P$ and the corresponding sequence $\sigma^B$, it holds that $B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) +
\beta$.
Combining this inequality with the inequalities between $P$ and $B$ and between \OFF and \OPT yields
\[
	2 \cdot \alpha \cdot P(\sigma^P) \leq 
	B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) + \beta
	 =\gamma \cdot 2 \cdot \alpha \cdot \OPT(\sigma^P) + \beta,
\]
and therefore $P$ is $\gamma$-competitive.
\end{proof}

>>>>

% for my small screen
%\setlength{\textwidth}{8cm}

@@

       \item lower and upper bound $k+1$ for two clusters, perfect partition and the optimal configuration reachable within one swap

@@
